---
title: "hw8_AvishaAvisha"
format: html
editor: visual
eval: true
---

```{r}
#| message: false

library(bayesplot)
library(tidyverse)
library(brms)
library(modelsummary)
library(dplyr)
library(posterior)
```

# Research Question

> Does the reaction time differ between participants with high versus low depression scores?

# Variables

-   \`ID' Participant id.
-   \`agentType': Participants reaction time in trust games was measured when they played against two different kinds of agents, represented by agent type variable.
-   'trustRTVec': Reaction time for each trial.
-   'DASS_Depr': Self-reported depression scores.

## Data Import

```{r}
data <- read.csv("merged_dat_trustgame_indvdfs.csv", row.names = NULL)

data$agentType <- as.factor(data$agentType)

data <- data[!is.na(data$trustRTVec) & !is.nan(data$trustRTVec) & !is.infinite(data$trustRTVec), ]
```

## Variable Summary

Table @tbl-summ-var shows the summary statistics of RTs by agent types.

```{r}
#| label: tbl-summ-var
#| tbl-cap: Descriptive statistics by groups
datasummary(trustRTVec*
                (N + Mean + SD + Min + Max + Histogram) ~ 
                factor(agentType),
            data = data)
```

Tibble @tbl-summ-var2 shows statistics for trustRTVec across ranges of depressionScores (0-21)

```{r}
#| label: tbl-summ-var2

data %>%
  group_by(DASS_Depr) %>%
  summarise(
    N = n(),
    Mean = mean(trustRTVec, na.rm = TRUE),
    SD = sd(trustRTVec, na.rm = TRUE),
    Min = min(trustRTVec, na.rm = TRUE),
    Max = max(trustRTVec, na.rm = TRUE)
  ) %>%
  print()
data %>%
  summarise(N = n(),
            Mean = mean(trustRTVec, na.rm = TRUE),
            SD = sd(trustRTVec, na.rm = TRUE),
            Min = min(trustRTVec, na.rm = TRUE),
            Max = max(trustRTVec, na.rm = TRUE))

```

# Model 1 using lognormal family

Let $Y_i$ = trustRTVec, $X_1$ = DASS_Depr, $X_2$ = agentType, $Z_i$ = ID

### Model:

$$
\begin{aligned}
Y_i & \sim \text{LogNormal}(\mu_i, \sigma) \\
\mu_i & = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 (X_1 \cdot X_2) + b_{0i} + b_{1i} X_2
\end{aligned}
$$

### Prior:

$$
\begin{aligned}
\beta_0 & \sim N(7, 2) \\
\beta_k & \sim N(0, 1), \quad k \in \{1, 2, 3\} \\
\sigma & \sim N^+(0, 3) \\
\end{aligned}
$$

## Analysis

Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1; total post-warmup draws = 4000 

```{r}
m1 <- brm(trustRTVec ~ DASS_Depr * agentType + (1 + agentType | ID), 
          family = lognormal(),
          data = data,
          prior <- c(
            prior(normal(7, 2), class = "Intercept"),      # Prior for intercept
            prior(normal(0, 1), class = "b"),              # Prior for fixed effects
            prior(normal(0, 3), class = "sigma")           # Prior for residual SD
          ),
          seed = 12345,
          file = "model_lognormal"
          )

```


# Results

Convergence checks

According to the rank histogram in @fig-rank-hist-fit below, the chains mixed well.

```{r}
#| label: fig-rank-hist-fit
#| fig-cap: Rank histogram of the posterior distributions of model parameters.

as_draws(m1) |>
  mcmc_rank_hist(pars = c(
    "b_Intercept",                          # Intercept
    "b_DASS_Depr",                          # Fixed effect for DASS_Depr
    "b_agentTypeGRGO",                      # Fixed effect for agentTypeGRGO
    "b_DASS_Depr:agentTypeGRGO",            # Interaction effect
    "sigma"                                 # Residual SD
  ))
```

@tbl-summ-fit shows the posterior distributions of the model parameters

```{r}
#| label: tbl-summ-fit
#| tbl-cap: Posterior summary of the model parameters.
summ_fit <- as_draws(m1) |>
    subset_draws(variable = c("b_Intercept", "b_DASS_Depr", "b_agentTypeGRGO", "b_DASS_Depr:agentTypeGRGO", "sigma")) |>
    summarise_draws()

# Display the summary as a table
knitr::kable(summ_fit, digits = 2)
```

@tbl-summ-fit2 shows the density plots of the posterior distributions of the model parameters.


```{r}
#| label: tbl-summ-fit2
#| tbl-cap: Density plot of the posterior distributions.
mcmc_areas(as_draws(m1), pars = c("b_Intercept", "b_DASS_Depr", "b_agentTypeGRGO", "b_DASS_Depr:agentTypeGRGO", "sigma"), prob = 0.95) # Show the 95% credible interval
```
# Interpretation

The results of the model suggest that the parameters related to **DASS Depression (b_DASS_Depr)** and the interaction between **DASS Depression** and **agentTypeGRGO** have negligible effects on the outcome variable **trustRTVec**, as their coefficients (mean = 0.00) are close to zero, with very small standard deviations. The **Intercept (b_Intercept)** is significantly greater than zero (mean = 6.49), indicating that the baseline value of the dependent variable is positive. The **agentTypeGRGO** parameter shows a slight negative effect (mean = -0.05), suggesting a small but significant decrease in trust response times for the **GRGO** group compared to the baseline. The **sigma** value (mean = 0.49) represents the residual standard deviation of the log-normal distribution, indicating some variability in the response times not explained by the model. All parameters show good model fit, with Rhat values close to 1 and high effective sample sizes (ESS), indicating reliable estimates. 

Overall these results suggest that **DASS Depression** scores do not significantly impact reaction times, and there is a minimal effect of **agentType** on the reaction times, although the baseline is positive. The interaction between depression and agent type does not seem to produce substantial changes in the response variable.

# Model 2 using skewnormal family

Let $Y_i$ = trustRTVec, $X_1$ = DASS_Depr, $X_2$ = agentType, $Z_i$ = ID

### Model:

$$
\begin{aligned}
Y_i & \sim \text{SkewNormal}(\mu_i, \sigma) \\
\mu_i & = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 (X_1 \cdot X_2) + b_{0i} + b_{1i} X_2
\end{aligned}
$$

### Prior:

$$
\begin{aligned}
\beta_0 & \sim N(7, 2) \\
\beta_k & \sim N(0, 1), \quad k \in \{1, 2, 3\} \\
\sigma & \sim N^+(0, 3) \\
\end{aligned}
$$

## Analysis

Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1; total post-warmup draws = 4000 

```{r}
m2 <- brm(trustRTVec ~ DASS_Depr * agentType + (1 + agentType | ID), 
          family = skew_normal(),
          data = data,
          prior <- c(
            prior(normal(7, 2), class = "Intercept"),      # Prior for intercept
            prior(normal(0, 1), class = "b"),              # Prior for fixed effects
            prior(normal(0, 3), class = "sigma")           # Prior for residual SD
          ),
          seed = 12345,
          file = "model_skewnormal"
          )

```


# Results

Convergence checks

According to the rank histogram in @fig-rank-hist-fit_m2 below, the chains mixed well.

```{r}
#| label: fig-rank-hist-fit_m2
#| fig-cap: Rank histogram of the posterior distributions of model parameters.

as_draws(m2) |>
  mcmc_rank_hist(pars = c(
    "b_Intercept",                          # Intercept
    "b_DASS_Depr",                          # Fixed effect for DASS_Depr
    "b_agentTypeGRGO",                      # Fixed effect for agentTypeGRGO
    "b_DASS_Depr:agentTypeGRGO",            # Interaction effect
    "sigma"                                 # Residual SD
  ))
```

@tbl-summ-fit_m2 shows the posterior distributions of the model parameters

```{r}
#| label: tbl-summ-fit_m2
#| tbl-cap: Posterior summary of the model parameters.

summ_fit <- as_draws(m2) |>
    subset_draws(variable = c("b_Intercept", "b_DASS_Depr", "b_agentTypeGRGO", "b_DASS_Depr:agentTypeGRGO", "sigma")) |>
    summarise_draws()

# Display the summary as a table
knitr::kable(summ_fit, digits = 2)
```

@tbl-summ-fit2_m2 shows the density plots of the posterior distributions of the model parameters.


```{r}
#| label: tbl-summ-fit2_m2
#| tbl-cap: Density plot of the posterior distributions.
#| 
mcmc_areas(as_draws(m2), pars = c("b_Intercept", "b_DASS_Depr", "b_agentTypeGRGO", "b_DASS_Depr:agentTypeGRGO", "sigma"), prob = 0.95) # Show the 95% credible interval
```

# Interpretation

The results from the **skew-normal** distribution model suggest that the parameters related to **DASS Depression (b_DASS_Depr)** and the interaction between **DASS Depression** and **agentTypeGRGO** have negligible effects on the outcome variable **trustRTVec**. The coefficients for these parameters are close to zero (mean = -0.09 for **DASS_Depr**, mean = -0.29 for the interaction), with large standard deviations, indicating substantial uncertainty and minimal impact on the outcome.

The **Intercept (b_Intercept)** is positive (mean = 10.10), with a wide range of possible values (from negative to large positive), suggesting that the baseline value of **trustRTVec** is on average higher but with significant variability.

The **agentTypeGRGO** parameter shows a slight negative effect (mean = -0.09), implying a small but uncertain decrease in trust response times for the **GRGO** group compared to the baseline. However, the wide credible intervals make this effect uncertain.

The **sigma** parameter (mean = 222.61) represents the residual standard deviation, indicating considerable unexplained variability in the response variable. This suggests that other unmeasured factors may be influencing **trustRTVec**, contributing to the variability that is not accounted for in the model.

All parameters show good model fit, with Rhat values close to 1 and high effective sample sizes (ESS), indicating reliable estimates despite the wide uncertainties in some parameters.

Overall, these results suggest that **DASS Depression** scores do not significantly impact **trustRTVec**, and the effect of **agentType** is small and uncertain. The interaction between **DASS Depression** and **agentType** does not seem to substantially affect the outcome variable. The model indicates significant unexplained variability, implying that other factors may be influencing **trustRTVec**. The results are similar to the log normal model m1. 